% Abstract — Manual FATEC 4.1.11 (versão do resumo em língua estrangeira).
% No original, o texto do abstract é apresentado em itálico.
\begin{resumo}[ABSTRACT]
\begin{otherlanguage}{english}
\itshape
The recent Brazilian economic scenario, characterized by macroeconomic
instability, high inflation, and a consequent increase in the Selic interest rate,
poses significant challenges for investors. Financial market volatility,
intensified by internal and external factors, hinders investment decision-making,
especially in assets such as stocks. Given this context and the increasing
digitalization of the financial sector, algorithmic trading emerges as a promising
alternative. This project proposes the development of a computational tool, using
the MQL5 language on the MetaTrader 5 platform, for analyzing trend and volatility
patterns in the Brazilian stock market, focusing on the PETR4 asset. The tool will
integrate statistical technical analyses, combining the use of Moving Averages for
trend identification, the ARIMA model for forecasting price time series, the GARCH
model for modeling the conditional volatility of ARIMA residuals, the RSI indicator
as a confirmation filter and Machine Learning for non-linear analysis. The central
hypothesis is that this integrated approach will allow the generation of trading
signals with greater effectiveness (potential accuracy and risk suitability)
compared to the isolated or incomplete use of these techniques. The study aims to
explore models with updated data, provide analyses that can support investment
strategies adapted to the high-interest-rate scenario, and investigate the
robustness of the combined models against market fluctuations, contributing to
knowledge in the field and offering a practical alternative for investors.

\vspace{\onelineskip}
\noindent\textbf{Keywords:} Algorithmic Trading. Statistical Models. ARIMA. GARCH.
Stock Market. Volatility. MQL5.
\end{otherlanguage}
\end{resumo}
